A Computational Account of Default-Mode Function\\ by Control Theory and Reinforcement Learning

### 2016-09-29 paper/paper.tex
more

Q-learning (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
\textit{greedy policy}
\begin{equation}
  \pi(s) = \pi_{\tilde{Q}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a),
\end{equation}
where $\tilde{Q}(s, a)$ is the current estimate of the state-action value table. In the sequel, except otherwise stated, the policies discussed will be of this form.
The hallmark of this trick wherein one only considers deterministic policies is the so-called
\textit{Bellman's equation}
\begin{equation}
  Q^\pi(s_t, a_t) = \mathbb E_{s_{t+1} \sim \rho(.|s_t)} [r(s_t,a_t) + \gamma Q^\pi(s_{t+1}, \pi(s_{t+1}))]
  \label{eq:bellman}
\end{equation}
which is a fixed-point equation for the $Q$-value operator of a deterministic policy $\pi$.
Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which
depend on the environment, and so can be learned off-policy by observing state-transitions
triggered by another behavioral policy, which can be stochastic.

\section{Q-learning with value approximation}
At state $s \in \mathcal S$, the DMN agent takes an action $a \in \mathcal A$ (for example,
$a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.) by sampling
from its current policy matrix, and collects a reward $r$ (measured on their internal currency of
worldly values ?), and the environment transitions to a new $s' \in \mathcal S$. At the end of
this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced: this represents an exemplar
behavior of the agent in such a situation, and is recorded in replay memory:
$\mathcal D \leftarrow \mathcal D \cup \{e\}$.
Now, replay consists in sampling (uniform or importance-weighted ?)  mini-batches of experiences
$e = (s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$ and trying to approximate
the would-be $Q$-value at $(s, a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
$r + \gamma \max_{a'} Q(s', a'|\theta^Q)$, with the output of a parametrized regression model $(s,a,r,s') \mapsto {Q}(s, a|\theta^Q)$, where $Q(.|\theta^Q)$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q$. Computing an optimal policy then corresponds to finding the parameters $\theta^Q$ which minimize the following loss function
\begin{equation}
  \mathcal L(\theta^Q)
  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q))^2]
  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - y)^2],
  \label{eq:oracle}
  \end{equation}
where
$ y = r + \gamma \max_{a'} Q(s', a'|\theta^Q) = r + \gamma Q(s', \pi(s')|\theta^Q)$ is the prediction target.
% While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.

@@

$${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
$\phi$ would represent a high-level hand-crafted representation of the state-action pairs into an appropriate feature space. Such are the techniques that have been proposed by [sutton et al. xyzw, Song et al. 2016 (NIPS), and Parr et al. 2003 (JMLR)].
A breakthrough alternative proposed by Google DeepMind [Minh et al. 2014, Silver et al 2015] is to learn this representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which are by far the current state-of-the-art in RL.

>>>>

\textit{Q-learning} (Watkins \& Dayan, 1992), a commonly used off-policy algorithm, uses the
\textit{greedy policy}
\begin{equation}
  \pi(s) = \pi_{\tilde{Q}}(s) = \argmax_{a \in \mathcal A}\tilde{Q}(s, a),
  \label{eq:qlearning}
\end{equation}
where $\tilde{Q}(s, a)$ is the current estimate of the state-action value table.
The hallmark of this trick wherein one only considers deterministic policies is the so-called
\textit{Bellman's equation}
\begin{equation}
  Q^\pi(s_t, a_t) = \mathbb E_{s_{t+1} \sim \rho(.|s_t)} [r(s_t,a_t) + \gamma Q^\pi(s_{t+1}, \pi(s_{t+1}))]
  \label{eq:bellman}
\end{equation}
which is a fixed-point equation for the $Q$-value operator of a deterministic policy $\pi$.
Notice that in the expectation in \eqref{eq:bellman}, the sampling is done over only things which
depend on the environment, and so can be learned off-policy by observing state-transitions
triggered by another behavioral policy, which can be stochastic.

\subsection{Q-learning with value approximation}
As already mentioned in the previous section, Q-learning optimizes over the class of
deterministic policies of the form \eqref{eq:qlearning}.
At state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$ (for example,
$a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.) by sampling
from its current policy matrix, and collects a reward $r$,
% (measured on their internal currency of worldly values ?),
and the environment transitions to a new $s' \in \mathcal S$. At the end of
this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced: this represents an exemplar
behavior of the agent in such a situation, and is recorded in replay memory:
$\mathcal D \leftarrow \mathcal D \cup \{e\}$.
Now, replay consists in sampling (uniform or importance-weighted ?)  mini-batches of experiences
$e = (s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$ and trying to
approximate
the would-be $Q$-value at $(s, a)$ as predicted by Bellman's equation \eqref{eq:bellman}, namely
$r + \gamma \max_{a'} Q(s', a'|\theta^Q)$, with the output of a parametrized regression model $(s,a,r,s') \mapsto {Q}(s, a|\theta^Q)$, where $Q(.|\theta^Q)$ is an approximator for the $Q$-value operator,  parameterized by $\theta^Q$. Computing an optimal policy then corresponds to finding the parameters $\theta^Q$ which minimize the following loss function
\begin{equation}
  \mathcal L(\theta^Q)
  % = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q))^2]
  = \mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - y)^2],
  \label{eq:oracle}
  \end{equation}
where
$ y = r + \gamma \max_{a'} Q(s', a'|\theta^Q) = r + \gamma Q(s', \pi(s')|\theta^Q)$ is the prediction target.
% While $y_t$ is also dependent on $\theta^Q$, this is typically ignored.

@@

$${Q}(s, a|\theta^Q) = \phi(s,a)^T\theta^Q.$$
$\phi$ would represent a high-level hand-crafted representation of the state-action pairs into an appropriate feature space. Such are the techniques that have been proposed by [sutton et al. xyzw, Song et al. 2016 (NIPS), and Parr et al. 2003 (JMLR)].
A breakthrough alternative proposed by Google DeepMind [Minh et al. 2014, Silver et al 2015] is to learn this representation using a deep neural net, leading to the so-called \textit{Deep Q-learning} family of methods which are by far the current state-of-the-art in RL. In such a case, $\theta^Q$ would correspond to ..
